[![R-CMD-check](https://github.com/reconhub/projections/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/reconhub/projections/actions/workflows/R-CMD-check.yaml)
[![codecov.io](https://codecov.io/github/reconhub/projections/coverage.svg?branch=master)](https://codecov.io/github/reconhub/projections?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/projections)](https://cran.r-project.org/package=projections)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.3923626.svg)](https://doi.org/10.5281/zenodo.3923626)


```{r options, echo = FALSE}
library(knitr)
opts_chunk$set(fig.width = 10, fig.height = 6, out.width = "80%")
```


# Welcome to the *projections* package!

This package uses data on *daily incidence*, the *serial interval* (time between
onsets of infectors and infectees) and the *reproduction number* to simulate
plausible epidemic trajectories and project future incidence. It relies on a
branching process where daily incidence follows a Poisson or a Negative Binomial
distribution governed by a force of infection computed (FOI) as:

$$
\lambda_t = \sum_{s = 1}^{t - 1} R_s y_s w(t - s)
$$

where:

* $w()$ is the probability mass function (PMF) of the serial interval
* $y_s$ is the incidence (by date of onset) at time $s$
* $R_s$ is the effective reproduction number (average number of secondary cases
  by infected case) at time $s$


Alternatively, the FOI can use the instantaneous reproduction number $R_t$
([Cori *et al.* 2013](https://academic.oup.com/aje/article/178/9/1505/89262))
instead of the effective reproduction number:

$$
\lambda_t = R_t \sum_{s = 1}^{t - 1} y_s w(t - s)
$$


## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("projections")
```

To benefit from the latest features and bug fixes, install the development,
*github* version of the package using:

```{r install2, eval = FALSE}
devtools::install_github("reconhub/projections")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:

- **`project`**: a function generating projections from an existing `incidence`
  object, a serial interval distribution, and a set of plausible reproduction
  numbers (*R*); returns a `projections` object; two models are implemented:
  Poisson, and Negative Binomial; both models can either use a constant
  distribution for *R*, or use time-varying distributions
  
- **`plot`/`print`**: plotting and printing methods for `projections` objects.

- **`summary`**: summary method for `projections` objects, deriving a range of
  summary statistics for each day of forecast

- **`get_dates`**: accessors for `projections` objects

- **`cumulate`**: cumulate predicted incidence over time

- **`as.data.frame`**: conversion from `projections` objects to `data.frame`

- **`[`**: subsetting operator for `projections` objects, permiting to specify
  which dates and simulations to retain; uses a syntax similar to matrices,
  i.e. `x[i, j]`, where `x` is the `projections` object, `i` a subset of dates,
  and `j` a subset of simulations

- **`subset`**: subset a `projections` object by specifying a time window

- **`build_projections`**: build a `projections` object from an input matrix and
  optional dates

- **`merge_projections`**: merge several `projections` objects into one, putting
  them on a common time frame; useful to combine runs from different simulations
  and/or perform model averaging

- **`merge_add_projections`**: puts several `projections` on the same time
  frame, and adds incidences to form a new object; objects having less
  simulations are recycled to match the largest number of simulations; useful to
  combine cases simulated from different processes

  



# Resources

## Worked example

### Simulated Ebola outbreak

In this example, we use the simulated Ebola outbreak `ebola_sim_clean` from the
`outbreaks` package to illustrate the package's functionalities. We will:

* first calculate case incidence
* create a serial interval distribution from known mean / standard deviations
  (e.g. taken from the literature)
* project case incidence
* summarise the resulting projections
* export results as `data.frame` for further processing, e.g. making custom
  plots using *ggplot2*
* showcase advanced handling of `projections` objects (merging/adding
  projections)


#### Computing case incidence

Here we use `incidence` (from the similarly named package) to calculate daily
counts of new cases by date of onset:

```{r onset}
library(outbreaks)
library(incidence)
library(ggplot2)

linelist <- ebola_sim_clean$linelist
i <- incidence(linelist$date_of_onset)
plot(i) +
  theme_bw() # full outbreak
plot(i[1:160]) +
  theme_bw() # first 160 days

```

#### Creating a serial interval

We create a serial interval distribution using `distcrete` (from the similarly
named package); we use published values of the Serial Interval for Ebola with a
mean of 15.3 days and a standard deviation of 9.3 days), to build a discretised
Gamma distribution. Because the Gamma implementation in R uses shape and scale
as parameters, we first need to convert the mean and coefficient of variation
into shape and scale, using `gamma_mucv2shapescale` from the *epitrix* package.

```{r si}

library(distcrete)
library(epitrix)
mu <- 15.3
sigma <- 9.3
cv <- sigma / mu
params <- gamma_mucv2shapescale(mu, cv)
params

si <- distcrete("gamma", shape = params$shape,
                scale = params$scale,
                interval = 1, w = 0.5)
si

si_df <- data.frame(t = 1:50,
                    p = si$d(1:50))
ggplot(si_df, aes(x = t, y = p)) +
  theme_bw() +
  geom_col() +
  labs(title = "Serial interval",
       x = "Days after onset",
       y = "Probability")

```


#### Projecting future incidence

We forecast future incidence based on the incidence data and the serial
interval, assuming a reproduction number of 1.5; for the sake of illustration,
we start use the first 100 days of data to determine the force of infection, and
derive forecasts for 30 days (in practice, forecasts can only be reliable for
short term predictions, typically 3 weeks maximum):

```{r predictions}

library(projections)
set.seed(1)
pred_1 <- project(i[1:100], R = 1.5, si = si, n_days = 30, n_sim = 1000)
pred_1
plot(pred_1) +
  theme_bw() # default plot
pred_1_cum <- cumulate(pred_1) # cumulative predictions
plot(pred_1_cum) +
  theme_bw() # plot cumulative predictions

```


Forecasts stored in a `projections` object can also be added to an `incidence`
plot using `add_projections`, which admits the same options as the `plot`
method. This function is best used with a pipe `%>%`:

```{r plot_with_incidence}

library(magrittr)
plot(i[20:160], alpha = 0.5) %>% 
  add_projections(pred_1, boxplots = FALSE, quantiles = c(0.025, 0.5)) +
  theme_bw()
 
```



#### Summarising forecasts

The `summary` function will summarise simulations using several statistics for
each day of the forecast, allowing the user to switch off some of the summaries,
and specify quantiles. Several options are illustrated below, but more
information will be found at `?summary.projections`:

``` {r summary}

## default summary
head(summary(pred_1))
tail(summary(pred_1))

## keeping only mean, min and max
head(summary(pred_1, sd = FALSE, quantiles = FALSE))

## using 10%, 50% and 90% quantiles
head(summary(pred_1, quantiles = c(0.1, 0.5, 0.9)))

```

To derive your own summary for each day, you can use `apply` with custom
functions; for instance, to calculate the geometric mean for each day:

```{r geom_mean}

## function to calculate geometric mean
geo_mean = function(x, na.rm = TRUE){
  exp(sum(log(x[x > 0]), na.rm = na.rm) / length(x))
}

## raw output
apply(pred_1, 1, geo_mean)

## with some formatting
temp <- apply(pred_1, 1, geo_mean)
data.frame(date = get_dates(pred_1),
           geometric_mean = apply(pred_1, 1, geo_mean),
           row.names = NULL)

```



#### Exporting results

The function `as.data.frame` can be handy for further processing of the
forecast. The argument `long` in particular will be handy for further processing
using `dplyr` or `ggplot2`, as it stores the 'simulation' as a 3rd columns,
which can be used for grouping and/or aesthetics. 

Here is an example with *ggplot2* to produce an alternative plot:

```{r plots}

df <- as.data.frame(pred_1, long = TRUE)
head(df)
ggplot(df, aes(x = date, y = incidence)) +
  theme_bw() +
  geom_jitter(alpha = 0.05, size = 4) +
  geom_smooth()

```



### Advanced handling

`projections` objects can also be combined in two ways:

1. **merge** different sets of simulations, using `merge_projections`; this can be
   useful e.g. for model averaging, where different models produce separate sets
   of forecasts which need combining
2. **add** forecasts from different simulation sets, using `+`, or
   `merge_add_projections`; this can be useful for simulating cases from
   different, complementary processes
   
We illustrate case 1, where we produce a second set of forecasts `pred_2` using
a different serial interval distribution, which we combine to `pred_1`. For the
sake of example, we use a made-up serial interval which is much shorter than the
one used in `pred_1`, with an average of 4 days, and a standard deviation of 2
days.

```{r other_si}

mu <- 4
sigma <- 2
cv <- sigma / mu
params <- gamma_mucv2shapescale(mu, cv)
params

si_short <- distcrete("gamma", shape = params$shape,
                      scale = params$scale,
                      interval = 1, w = 0.5)
si_short

si_short_df <- data.frame(t = 1:20,
                          p = si_short$d(1:20))
ggplot(si_short_df, aes(x = t, y = p)) +
  theme_bw() +
  geom_col() +
  labs(title = "Other serial interval",
       x = "Days after onset",
       y = "Probability")

```

We now use this serial interval to produce a second set of forecasts. We compare
them to the initial one, and the combined forecasts:

```{r pred_2}

set.seed(1)
pred_2 <- project(i[1:100], R = 1.5, si = si_short, n_days = 30, n_sim = 1000)
pred_2 # 1000 simulations
plot(pred_2) +
  theme_bw() # default plot

## combine the objects; note that any number of objects can be combined
pred_combined <- merge_projections(list(pred_1, pred_2))
pred_combined # 2000 simulations

list_plots <- list(
  plot(pred_1) + theme_bw() + labs(title = "Forecast with initial SI"),
  plot(pred_2,) + theme_bw() + labs(title = "Forecast with short SI"),
  plot(pred_combined) + theme_bw() + labs(title = "Combined forecasts")
)

library(cowplot)
plot_grid(plotlist = list_plots,
          ncol = 1)

```


To illustrate case 2 (not only merging, but adding `projections` objects), we
artificially split the dataset by hospitals, derive separate forecasts for each,
and add forecasts of two hospitals in the example below.

```{r adding_forecasts}

## calculate incidence by hospital
i_hosp <- incidence(linelist$date_of_onset, groups = linelist$hospital)
plot(i_hosp) +
  theme_bw() +
  theme(legend.position = "bottom")


## derive predictions for each hospital
n_groups <- ncol(get_counts(i_hosp))

pred_hosp <- lapply(
  seq_len(n_groups),
  function(j)
    project(i_hosp[1:100, j],
            R = 1.5,
            si = si,
            n_days = 60,
            n_sim = 500))
names(pred_hosp) <- colnames(get_counts(i_hosp))


## we combine forecasts for Connaught and Rokupa hospitals
pred_connaught_rokupa <- pred_hosp$`Connaught Hospital` + pred_hosp$`Rokupa Hospital`


list_plots <- list(
  plot(pred_hosp$`Connaught Hospital`) +
    theme_bw() +
    ylim(c(0, 30)) +
    labs(title = "Connaught hospital"),
  plot(pred_hosp$`Rokupa Hospital`) +
    theme_bw() +
    ylim(c(0, 30)) +
    labs(title = "Rokupa hospital"),
  plot(pred_connaught_rokupa) +
    theme_bw() +
    ylim(c(0, 30)) +
    labs(title = "Connaught + Rokupa")
)

plot_grid(plotlist = list_plots,
          ncol = 1)

```

Note that to add more than 2 `projections` objects, one can use
`merge_add_projections`. Also note that the `+` operator can also be used with a
`numeric`, e.g. for adding an offset. For instance:

```{r offset}

plot(pred_connaught_rokupa + 1000) +
  theme_bw() +
  labs(title = "Connaught + Rokupa with 1000 added cases")

```




## Vignettes

*projections* does not currently have a dedicated vignette. This `README` will
eventually be converted into separate vignettes.


## Websites

A dedicated website can be found at:
[http://www.repidemicsconsortium.org/projections](http://www.repidemicsconsortium.org/projections).






## Getting help online

Bug reports and feature requests should be posted on *github* using the
[*issue*](http://github.com/reconhub/projections/issues) system. All other
questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via [pull
requests](https://github.com/reconhub/projections/pulls).

Please note that this project is released with a [Contributor Code of
Conduct](CONDUCT.md). By participating in this project you agree to abide by its
terms.

