[![Build Status](https://travis-ci.org/reconhub/projections.svg?branch=master)](https://travis-ci.org/reconhub/projections)
[![Build status](https://ci.appveyor.com/api/projects/status/265h2el4y9popan9/branch/master?svg=true)](https://ci.appveyor.com/project/thibautjombart/projections/branch/master)
[![codecov.io](https://codecov.io/github/reconhub/projections/coverage.svg?branch=master)](https://codecov.io/github/reconhub/projections?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/projections)](https://cran.r-project.org/package=projections)


```{r options, echo = FALSE}
library(knitr)
opts_chunk$set(fig.width = 10, fig.height = 6, out.width = "80%")
```


# Welcome to the *projections* package!

This package uses data on *daily incidence*, the *serial interval* (time between
onsets of infectors and infectees) and the *reproduction number* to simulate
plausible epidemic trajectories and project future incidence. It relies on a
branching process where daily incidence follows a Poisson process determined by
a daily infectiousness, computed as:

$$
\lambda_t = \sum_{s = 1}^{t - 1} y_s w(t - s)
$$

where $w()$ is the probability mass function (PMF) of the serial interval, and
$y_s$ is the incidence at time $s$.


## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("projections")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval = FALSE}
devtools::install_github("reconhub/projections")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:

- **`project`**: a function generating projections from an existing *incidence*
  object, a serial interval distribution, and a set of plausible reproduction
  numbers ($R$); returns a `projections` object.
  
- **`plot`/`print`**: plotting and printing methods for `projections` objects.

- **`get_dates`/`get_incidence`**: accessors for `projections` objects.

- **`as.data.frame`**: conversion from `projections` objects to `data.frame`.


# Resources

## Worked example

In the following, we project future incidence using a simulate Ebola outbreak
distributed in the package `outbreaks`:
```{r onset}
library(outbreaks)
library(incidence)

onset <- ebola_sim$linelist$date_of_onset
i <- incidence(onset)
plot(i) # full outbreak
plot(i[1:160]) # first 160 days
```

We create a serial interval distribution using `distcrete`:
```{r si}
library(distcrete)
library(epitrix)
mu <- 15.3
sigma <- 9.3
cv <- sigma / mu
params <- gamma_mucv2shapescale(mu, cv)
params

si <- distcrete("gamma", shape = params$shape,
                scale = params$scale,
                interval = 1, w = 0)
si

plot(1:50, si$d(1:50), type = "h", lwd = 3, col = "navy",
     main = "Serial interval", xlab = "Days after onset",
     ylab = "Relative infectiousness")
```


We predict future incidence based on these data, assuming a reproduction number
of 1.5, from day 100 and for 60 days:

```{r predictions}
library(projections)
set.seed(1)
pred <- project(i[1:100], R = 1.5, si = si, n_days = 60)
pred
plot(pred) # default plot of quantiles
apply(pred, 1, mean) # average prediction per day
apply(pred, 1, range) # range across simulations
```

An alternative representation of the outcomes:
```{r plots}
library(ggplot2)
df <- as.data.frame(pred, long = TRUE)
head(df)
p <- ggplot(df, aes(x = date, y = incidence)) +
  geom_jitter(alpha = .3) + geom_smooth()
p
```


Predictions can also be added to the epicurve:
```{r plot_with_incidence}
library(magrittr)

plot(i[20:160]) %>% add_projections(pred)
 
```




## Vignettes

*projections* does not currently have a dedicated vignette; instead, it is
illustrated in conjunction with `earlyR` on [this
vignette](http://www.repidemicsconsortium.org/earlyR/articles/earlyR.html).


## Websites

A dedicated website can be found at:
[http://www.repidemicsconsortium.org/projections](http://www.repidemicsconsortium.org/projections).






## Getting help online

Bug reports and feature requests should be posted on *github* using the
[*issue*](http://github.com/reconhub/projections/issues) system. All other
questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via [pull
requests](https://github.com/reconhub/projections/pulls).

Please note that this project is released with a [Contributor Code of
Conduct](CONDUCT.md). By participating in this project you agree to abide by its
terms.

