[![Build Status](https://travis-ci.org/reconhub/projections.svg?branch=master)](https://travis-ci.org/reconhub/projections)
[![Build status](https://ci.appveyor.com/api/projects/status/265h2el4y9popan9/branch/master?svg=true)](https://ci.appveyor.com/project/thibautjombart/projections/branch/master)
[![codecov.io](https://codecov.io/github/reconhub/projections/coverage.svg?branch=master)](https://codecov.io/github/reconhub/projections?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/projections)](https://cran.r-project.org/package=projections)


```{r options, echo = FALSE}
opts_chunk$set(fig.width = 6, fig.height = 4))
```

# Welcome to the *projections* package!

This package uses data on *daily incidence*, the *serial interval* (time between
onsets of infectors and infectees) and the *reproduction number* to simulate
plausible epidemic trajectories and project future incidence. It relies on a
branching process where daily incidence follows a Poisson process determined by
a daily infectiousness, computed as:

$$
\lambda_t = \sum_{s = 1}^{t - 1} y_s w(t - s)
$$

where $w()$ is the probability mass function (PMF) of the serial interval, and
$y_s$ is the incidence at time $s$.


## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("projections")
```

To benefit from the latest features and bug fixes, install the development, *github* version of the package using:
```{r install2, eval = FALSE}
devtools::install_github("reconhub/projections")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:

- **`project`**: a function generating projections from an existing *incidence*
  object, a serial interval distribution, and a set of plausible reproduction
  numbers ($R$); returns a `projections` object.
  
- **`plot`/`print`**: plotting and printing methods for `projections` objects.

- **`get_dates`/`get_incidence`**: accessors for `projections` objects.

- **`as.data.frame`**: conversion from `projections` objects to `data.frame`.


# Resources

## Worked example

In the following, we simulate a small outbreak with the following dates of
onset:
```{r onset}
onset <- as.Date(c("2017-12-01", "2017-12-04", "2017-12-05", "2017-12-05",
                 "2017-12-05", "2017-12-06", "2017-12-08"))
                
```

The package incidence is used to generate an epicurve:

```{r epicurve}
library(incidence)
i <- incidence(onset)
i
plot(i)
```

Let us assume the following serial interval distribution with a mean of 2.8 days
and a CV of +/- 0.4 days (in practice, these values would likely come from the literature):

```{r interval}
library(distcrete)
library(epitrix)
mu <- 2.8
cv <- 0.4
params <- gamma_mucv2shapescale(mu, cv)
params
si <- distcrete("gamma", shape = params$shape, scale = params$scale, interval = 1, w = 0)
si
plot(si$d, xlim = c(0,10), type = "h", lwd = 3, col = "navy",
     main = "Serial interval", xlab = "Days after onset", ylab = "Relative infectiousness")
```

We also need to estimate transmissibility. For this, we use the function `get_R`
from the `earlyR` package:

```{r reproduction-number}
library(earlyR)
R <- get_R(i, si = si) # ML estimation
R_samp <- sample_R(R, 100) # 100 plausible values of R
```



We can now predict future incidence based on these data:

```{r predictions}
library(projections)
set.seed(1)
pred <- project(i, R = R_samp, si = si, n_days = 14)
pred
plot(pred) # median and 95% prediction
apply(pred, 1, mean) # average prediction per day
apply(pred, 1, range) # range across simulations
```

An alternative representation of the outcomes:
```{r plots}
library(ggplot2)
df <- as.data.frame(pred, long = TRUE)
head(df)
p <- ggplot(df, aes(x = date, y = incidence)) + geom_jitter(alpha = .3) + geom_smooth()
p
```



## Vignettes

*projections* does not currently have a dedicated vignette; instead, it is illustrated in conjunction with `earlyR` on [this vignette](http://www.repidemicsconsortium.org/earlyR/articles/earlyR.html).


## Websites

A dedicated website can be found at:
[http://www.repidemicsconsortium.org/projections](http://www.repidemicsconsortium.org/projections).






## Getting help online

Bug reports and feature requests should be posted on *github* using the [*issue*](http://github.com/reconhub/projections/issues) system. All other questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via [pull requests](https://github.com/reconhub/projections/pulls).

Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

