[![Build Status](https://travis-ci.org/reconhub/projections.svg?branch=master)](https://travis-ci.org/reconhub/projections)
[![Build status](https://ci.appveyor.com/api/projects/status/265h2el4y9popan9/branch/master?svg=true)](https://ci.appveyor.com/project/thibautjombart/projections/branch/master)
[![codecov.io](https://codecov.io/github/reconhub/projections/coverage.svg?branch=master)](https://codecov.io/github/reconhub/projections?branch=master)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/projections)](https://cran.r-project.org/package=projections)


```{r options, echo = FALSE}
library(knitr)
opts_chunk$set(fig.width = 10, fig.height = 6, out.width = "80%")
```


# Welcome to the *projections* package!

This package uses data on *daily incidence*, the *serial interval* (time between
onsets of infectors and infectees) and the *reproduction number* to simulate
plausible epidemic trajectories and project future incidence. It relies on a
branching process where daily incidence follows a Poisson process determined by
a daily infectiousness, computed as:

$$
\lambda_t = \sum_{s = 1}^{t - 1} R_s y_s w(t - s)
$$

where:

* $w()$ is the probability mass function (PMF) of the serial interval
* $y_s$ is the incidence (by date of onset) at time $s$
* $R_s$ is the effective reproduction number (average number of secondary cases
  by infected case) at time $s$


## Installing the package

To install the current stable, CRAN version of the package, type:
```{r install, eval = FALSE}
install.packages("projections")
```

To benefit from the latest features and bug fixes, install the development,
*github* version of the package using:

```{r install2, eval = FALSE}
devtools::install_github("reconhub/projections")
```

Note that this requires the package *devtools* installed.


# What does it do?

The main features of the package include:

- **`project`**: a function generating projections from an existing `incidence`
  object, a serial interval distribution, and a set of plausible reproduction
  numbers (*R*); returns a `projections` object; two models are implemented:
  Poisson, and Negative Binomial; both models can either use a constant
  distribution for *R*, or use time-varying distributions
  
- **`plot`/`print`**: plotting and printing methods for `projections` objects.

- **`get_dates`**: accessors for `projections` objects.

- **`cumulate`**: cumulate predicted incidence over time.

- **`as.data.frame`**: conversion from `projections` objects to `data.frame`.

- **`[`**: subsetting operator for `projections` objects, permiting to specify
  which dates and simulations to retain; uses a syntax similar to matrices,
  i.e. `x[i, j]`, where `x` is the `projections` object, `i` a subset of dates,
  and `j` a subset of simulations.

- **`subset`**: subset a `projections` object by specifying a time window.

- **`build_projections`**: build a `projections` object from an input matrix and
  optional dates.

- **`merge_projections`**: merge several `projections` objects into one, putting
  them on a common time frame; useful to combine runs from different simulations
  and/or perform model averaging

- **`merge_add_projections`**: puts several `projections` on the same time
  frame, and adds incidences to form a new object; objects having less
  simulations are recycled to match the largest number of simulations; useful to
  combine cases simulated from different processes

  



# Resources

## Worked example

### Simulated Ebola outbreak

In this example, we use the simulated Ebola outbreak `ebola_sim_clean` from the
`outbreaks` package to illustrate the package's functionalities. We will:

* first calculate case incidence
* create a serial interval distribution from known mean / standard deviations
  (e.g. taken from the literature)
* project case incidence
* summarise the resulting projections
* export results as `data.frame` for further processing, e.g. making custom
  plots using *ggplot2*
* showcase advanced handling of `projections` objects (merging/adding
  projections)


#### Compute case incidence.

Here we use `incidence` to calculate daily counts of new cases by date of onset:

```{r onset}
library(outbreaks)
library(incidence)
library(ggplot2)

linelist <- ebola_sim_clean$linelist
i <- incidence(linelist$date_of_onset)
plot(i) +
  theme_bw() # full outbreak
plot(i[1:160]) +
  theme_bw() # first 160 days

```

We create a serial interval distribution using `distcrete`:
```{r si}
library(distcrete)
library(epitrix)
mu <- 15.3
sigma <- 9.3
cv <- sigma / mu
params <- gamma_mucv2shapescale(mu, cv)
params

si <- distcrete("gamma", shape = params$shape,
                scale = params$scale,
                interval = 1, w = 0.5)
si

si_df <- data.frame(t = 1:50,
                    p = si$d(1:50))
ggplot(si_df, aes(x = t, y = p)) +
  theme_bw() +
  geom_col() +
  labs(title = "Serial interval",
       x = "Days after onset",
       y = "Probability")

```


We predict future incidence based on these data, assuming a reproduction number
of 1.5, from day 100 and for 60 days:

```{r predictions}
library(projections)
set.seed(1)
pred_1 <- project(i[1:100], R = 1.5, si = si, n_days = 60, n_sim = 1000)
pred_1
plot(pred_1) +
  theme_bw() # default plot
pred_1_cum <- cumulate(pred_1) # cumulative predictions
plot(pred_1_cum) +
  theme_bw() # plot cumulative predictions

apply(pred_1, 1, mean) # average prediction per day
apply(pred_1, 1, quantile, c(0.025, 0.25, 0.5, 0.75, 0.975)) # quantiles

```

Note that you 
An alternative representation of the outcomes:

```{r plots}

df <- as.data.frame(pred_1, long = TRUE)
head(df)
p <- ggplot(df, aes(x = date, y = incidence)) +
  geom_jitter(alpha = .3) + geom_smooth()
p

```


Predictions can also be added to the epicurve:
```{r plot_with_incidence}
library(magrittr)

plot(i[20:160]) %>% add_projections(pred, boxplots = FALSE)
 
```



### Combining models

What if we were uncertain about the serial interval, and wanted to use, say, 3
different serial intervals, and study the impact of this uncertainty on our
predictions? One way to approach this is to create 3 sets of simulations, and
merge the resulting objects

```{r }

## get incidence by hospital
i_hosp <- incidence(linelist$date_of_onset, groups = linelist$hospital)
plot(i_hosp) + theme_bw()


## derive predictions for each hospital
n_groups <- ncol(get_counts(i_hosp))

pred_hosp <- lapply(
  seq_len(n_groups),
  function(j)
    project(i_hosp[1:100, j],
            R = 1.5,
            si = si,
            n_days = 60,
            n_sim = 500))
names(pred_hosp) <- colnames(get_counts(i_hosp))
cowplot::plot_grid(plotlist = lapply(pred_hosp, plot), labels = names(pred_hosp))

## combine predictions only for known hospitals
pred_merged <- merge_add_projections(pred_hosp)

## note: we can also use `+`; for instance to combine predictions for Connaught
## and Rokupa hospitals: 
pred_connaugt_rokupa <- pred_hosp$`Connaught Hospital` + pred_hosp$`Rokupa Hospital`
cowplot::plot_grid(
  plot(pred_hosp$`Connaught Hospital`),
  plot(pred_hosp$`Rokupa Hospital`),
  plot(pred_connaugt_rokupa),
  ncol = 3,
  labels = c("Connaught", "Rokupa", "Connaught + Rokupa"))

```






## Vignettes

*projections* does not currently have a dedicated vignette; instead, it is
illustrated in conjunction with `earlyR` on [this
vignette](http://www.repidemicsconsortium.org/earlyR/articles/earlyR.html).


## Websites

A dedicated website can be found at:
[http://www.repidemicsconsortium.org/projections](http://www.repidemicsconsortium.org/projections).






## Getting help online

Bug reports and feature requests should be posted on *github* using the
[*issue*](http://github.com/reconhub/projections/issues) system. All other
questions should be posted on the **RECON forum**: <br>
[http://www.repidemicsconsortium.org/forum/](http://www.repidemicsconsortium.org/forum/)

Contributions are welcome via [pull
requests](https://github.com/reconhub/projections/pulls).

Please note that this project is released with a [Contributor Code of
Conduct](CONDUCT.md). By participating in this project you agree to abide by its
terms.

